Carbonate chemistry at Ny-Ålesund

### 2023-01-02 awipev-CO2.Rmd
sorting out the mess in stratification (in progress)

```{r, echo=FALSE, message= FALSE, warning=FALSE}
# we look at the distribution of data
depth_labels <- c("0-2", "2-4", "4-6", "6-8", ">8")
# we look at salinity all depths
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = sal_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Salinity vs depth", x = "Depth (m)", y= "Salinity") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$sal_insitu_ctd ~ tmp$depth_group * tmp$montht))

# we look at salinity 0 to 4 m vs more than 8 m
tmp <- d_all %>%
  dplyr::select(montht, sal_insitu_ctd, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(
    case_when(
      depth_group == '0-2' | depth_group == '2-4' ~ '0-4',
      depth_group == '>8' ~ '8-12'
    )
  )) %>%
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>%
  dplyr::group_by(depth_group2, montht)
tmp_delta <- tidyr::drop_na(tmp) %>%
  group_by(depth_group2, montht) %>%
  dplyr::summarise(sal_med = median(sal_insitu_ctd, na.rm = TRUE)) %>%
  tidyr::pivot_wider(names_from = depth_group2, values_from = sal_med) %>%
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>%
  dplyr::mutate(sal_delta = median0_4m - median8_12m)
kable(tmp_delta, digits = 2, caption = "Median salinity with (median0_4m - median8_12m)")

@@

```{r, echo=FALSE, message= FALSE, warning=FALSE}
# we look at temperature all depths
tmp <- d_all %>% 
  dplyr::select(montht, temp_insitu_ctd, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = temp_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Temperature vs depth", x = "Depth (m)", y= "Temperature") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$temp_insitu_ctd ~ tmp$depth_group * tmp$montht))
# we look at temperature 0 to 4 m vs more than 8 m
tmp <- d_all %>%
  dplyr::select(montht, temp_insitu_ctd, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(
    case_when(
      depth_group == '0-2' | depth_group == '2-4' ~ '0-4',
      depth_group == '>8' ~ '8-12'
    )
  )) %>%
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>%
  dplyr::group_by(depth_group2, montht)
tmp_delta <- tmp %>%
  group_by(depth_group2, montht) %>%
  dplyr::summarise(temp_med = median(temp_insitu_ctd, na.rm = TRUE)) %>%
  tidyr::pivot_wider(names_from = depth_group2, values_from = temp_med) %>%
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>%
  dplyr::mutate(temp_delta = median0_4m - median8_12m)
kable(tmp_delta, digits = 2, caption = "Median temperature with (median0_4m - median8_12m)")
p <- tmp %>% 
  tidyr::drop_na() %>% 
  ggplot(aes(x=depth_group2, y = temp_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Temperature vs depth", x = "Depth (m)", y= "Temperature") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
temp_delta_plt <- tmp_delta %>% 
  #tidyr::drop_na() %>% 
  ggplot(aes(x = montht, y = temp_delta)) +
  #scale_x_discrete(expand = c(0, 0)) +
  scale_x_discrete("Month") +
  scale_y_continuous("Temperature", limits = c(-2.1, 0.5), expand = c(0, 0)) +
  geom_col(fill="blue") +
  geom_hline(yintercept = 0) +
  annotate("text", x = 2, y = 0.3, parse = TRUE, label = "T['surface'] > T['deep']") +
  annotate("text", x = 2, y = -1.8, parse = TRUE, label = "T['surface'] <  T['deep']") +
  Mytheme(size_labs = 12) +
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())
print(temp_delta_plt)
```

@@

```{r, echo=FALSE, message= FALSE, warning=FALSE}
# we look at density at all depths
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, temp_insitu_ctd, depth_group, pressure_insitu_ctd) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  dplyr::mutate(density = seacarb::rho(S = sal_insitu_ctd, T = temp_insitu_ctd, P = pressure_insitu_ctd)) %>% 
  group_by(depth_group, montht)
p <- tidyr::drop_na(tmp) %>% 
  ggplot(aes(x=as.factor(depth_group), y = sal_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Density vs depth", x = "Depth (m)", y= "Density") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$density ~ tmp$depth_group * tmp$montht))

# we look at density 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, temp_insitu_ctd, depth_group, pressure_insitu_ctd) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  dplyr::mutate(depth_group2 = as.factor(case_when(depth_group == '0-2' | depth_group == '2-4' ~ '0-4', 
                                      depth_group == '>8' ~ '8-12')), 
                density = seacarb::rho(S = sal_insitu_ctd, T = temp_insitu_ctd, 
                                       P = pressure_insitu_ctd)) %>% 
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>% 
  dplyr::group_by(depth_group2, montht)

@@

```{r, echo=FALSE, message= FALSE, warning=FALSE}
# we look at pHint_tot_sf at all depths
tmp <- d_all %>% 
  dplyr::select(montht, pHint_tot_sf, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x = as.factor(depth_group), y = pHint_tot_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pHint_tot_sf vs depth", x = "Depth (m)", y= "pH (units)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pHint_tot_sf ~ tmp$depth_group * tmp$montht))

@@

tmp_delta <- tmp %>%
  group_by(depth_group2, montht) %>%
  dplyr::summarise(pH_tot_sf_med = median(pHint_tot_sf, na.rm = TRUE)) %>%
  tidyr::pivot_wider(names_from = depth_group2, values_from = pH_tot_sf_med) %>%
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>%   dplyr::mutate(pH_tot_sf_med_delta = median0_4m - median8_12m)
kable(tmp_delta, digits = 3, caption = "Median seafet pH with (median0_4m - median8_12m)")

@@

# we look at pHint_tot_sf normalized at 4°C at all depths
tmp <- d_all %>% 
  dplyr::select(montht, pHint_tot_sf_temp_4, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = pHint_tot_sf_temp_4)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pHint_tot_sf at 4°C vs depth", x = "Depth (m)", y= "pH at 4°C (units)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pHint_tot_sf_temp_4 ~ tmp$depth_group * tmp$montht))

# we look at pHint_tot_sf normalized at 4°C  0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, pHint_tot_sf_temp_4, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(case_when(depth_group == '0-2' | depth_group == '2-4' ~ '0-4',
                                                   depth_group == '>8' ~ '8-12'))) %>% 
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>% 
  dplyr::group_by(depth_group2, montht)
summary(aov(tmp$pHint_tot_sf_temp_4 ~ tmp$depth_group2 * tmp$montht))

@@

# we look at pHint_tot_sf normalized at 4°C at all depths
tmp <- d_all %>% 
  dplyr::select(montht, pHint_tot_sf_temp_4, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = pHint_tot_sf_temp_4)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pHint_tot_sf at 4°C vs depth", x = "Depth (m)", y= "pH at 4°C (units)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pHint_tot_sf_temp_4 ~ tmp$depth_group * tmp$montht))

# we look at pHint_tot_sf normalized at 4°C  0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, pHint_tot_sf_temp_4, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(case_when(depth_group == '0-2' | depth_group == '2-4' ~ '0-4',
                                                   depth_group == '>8' ~ '8-12'))) %>% 
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>% 
  dplyr::group_by(depth_group2, montht)
summary(aov(tmp$pHint_tot_sf_temp_4 ~ tmp$depth_group2 * tmp$montht))

@@

```{r, echo=FALSE, message= FALSE, warning=FALSE}
# we look at pCO2 NOT normalized at 4°C at all depths
# calculate pCO2 from pH and salinity-derived AT
tmp <- d_all %>%
  dplyr::select(datetime,
                montht,
                depth_group,
                depth,
                pHint_tot_sf,
                sal_insitu_ctd,
                temp_insitu_ctd) %>%
  drop_na() %>%
  dplyr::mutate(
    at_remos = (intercept_s_at + slope_s_at * sal_insitu_ctd) * 1e-6,
    pco2_remos_ctd_temp = seacarb::carb(
      flag = 8,
      var1 = pHint_tot_sf,
      var2 = at_remos,
      S = sal_insitu_ctd,
      T = temp_insitu_ctd,
      Patm = 1,
      P = depth / 10,
      Pt = 0,
      Sit = 0,
      k1k2 = "l",
      kf = "dg",
      ks = "d",
      pHscale = "T",
      b = "u74"
    )$pCO2
  )
# add at_remos and pco2_remos_ctd_temp to d_all
d_all <-
  dplyr::left_join(d_all, dplyr::select(tmp, datetime, at_remos, pco2_remos_ctd_temp), by = "datetime") %>%
  dplyr::select(sort(current_vars()))
tmp <- tmp %>% 
  dplyr::select(montht, pco2_remos_ctd_temp, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = pco2_remos_ctd_temp)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pco2_sf vs depth", x = "Depth (m)", y= "fCO2 (uatm)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pco2_remos_ctd_temp ~ tmp$depth_group * tmp$montht))

# we look at pCO2 NOT normalized at 4°C 0 to 4 m vs more than 8 m
tmp <- tmp %>%
  dplyr::select(montht, pco2_remos_ctd_temp, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(
    case_when(
      depth_group == '0-2' | depth_group == '2-4' ~ '0-4',
      depth_group == '>8' ~ '8-12'
    )
  )) %>%
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>%
  dplyr::group_by(depth_group2, montht)
summary(aov(tmp$pco2_remos_ctd_temp ~ tmp$depth_group2 * tmp$montht))

tmp_delta <- tmp %>%
  group_by(depth_group2, montht) %>%
  dplyr::summarise(pco2_remos_ctd_temp_med = median(pco2_remos_ctd_temp, na.rm = TRUE)) %>%
  tidyr::pivot_wider(names_from = depth_group2, values_from = pco2_remos_ctd_temp_med) %>%
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>%
  dplyr::mutate(pco2_remos_ctd_temp_med_delta = median0_4m - median8_12m) %>%
  # this is to bring June back...
  rbind(
    data.frame(
      montht = "Jun",
      median8_12m = NA,
      median0_4m = NA,
      pco2_remos_ctd_temp_med_delta = NA
    )
  ) %>%
  dplyr::mutate(montht = factor(montht, levels = month.abb)) %>%
  dplyr::arrange(montht)
kable(tmp_delta, digits = 3, caption = "Median REMOS pCO2 (median0_4m - median8_12m)")

@@

tmp <- tmp %>% 
  dplyr::select(montht, pco2_remos_4, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  group_by(depth_group, montht)
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = pco2_remos_4)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pco2_remos at 4°C vs depth", x = "Depth (m)", y= "pCO2 (uatm)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pco2_remos_4 ~ tmp$depth_group * tmp$montht))

>>>>

```{r, echo=FALSE, message= FALSE, warning=FALSE}
# we look at the distribution of data
depth_labels <- c("0-2", "2-4", "4-6", "6-8", ">8")
# we look at salinity all depths
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, depth_group) %>%
  dplyr::filter(depth_group != "NA") %>% 
  dplyr::mutate(depth_group = factor(depth_group, levels = depth_labels))
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = sal_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Salinity vs depth", x = "Depth (m)", y= "Salinity") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$sal_insitu_ctd ~ tmp$depth_group * tmp$montht))

# we look at salinity 0 to 4 m vs more than 8 m
tmp <- d_all %>%
  dplyr::select(montht, sal_insitu_ctd, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(
    case_when(
      depth_group == '0-2' | depth_group == '2-4' ~ '0-4',
      depth_group == '>8' ~ '8-12'
    )
  )) %>%
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') #%>%
tmp_delta <- tidyr::drop_na(tmp) %>%
  group_by(depth_group2, montht) %>%
  dplyr::summarise(sal_med = median(sal_insitu_ctd, na.rm = TRUE)) %>%
  tidyr::pivot_wider(names_from = depth_group2, values_from = sal_med) %>%
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>%
  dplyr::mutate(sal_delta = median0_4m - median8_12m)
kable(tmp_delta, digits = 2, caption = "Median salinity with (median0_4m - median8_12m)")

@@

```{r, echo=FALSE, message= FALSE, warning=FALSE}
# we look at temperature all depths
tmp <- d_all %>% 
  dplyr::select(montht, temp_insitu_ctd, depth_group) %>%
  dplyr::filter(depth_group != "NA") %>% 
  dplyr::mutate(depth_group = factor(depth_group, levels = depth_labels))
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = temp_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Temperature vs depth", x = "Depth (m)", y= "Temperature") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$temp_insitu_ctd ~ tmp$depth_group * tmp$montht))
# we look at temperature 0 to 4 m vs more than 8 m
tmp <- d_all %>%
  dplyr::select(montht, temp_insitu_ctd, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(
    case_when(
      depth_group == '0-2' | depth_group == '2-4' ~ '0-4',
      depth_group == '>8' ~ '8-12'
    )
  )) %>%
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>%
  dplyr::group_by(depth_group2, montht)
tmp_delta <- tmp %>%
  #group_by(depth_group2, montht) %>%
  dplyr::summarise(temp_med = median(temp_insitu_ctd, na.rm = TRUE)) %>%
  tidyr::pivot_wider(names_from = depth_group2, values_from = temp_med) %>%
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>%
  dplyr::mutate(temp_delta = median0_4m - median8_12m)
kable(tmp_delta, digits = 2, caption = "Median temperature with (median0_4m - median8_12m)")
p <- tmp %>% 
  tidyr::drop_na() %>% 
  ggplot(aes(x=depth_group2, y = temp_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Temperature vs depth", x = "Depth (m)", y= "Temperature") +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
temp_delta_plt <- tmp_delta %>% 
  ggplot(aes(x = montht, y = temp_delta)) +
  scale_x_discrete("Month") +
  scale_y_continuous("Temperature", limits = c(-2.1, 0.5), expand = c(0, 0)) +
  geom_col(fill="blue") +
  geom_hline(yintercept = 0) +
  annotate("text", x = 2, y = 0.3, parse = TRUE, label = "T['surface'] > T['deep']") +
  annotate("text", x = 2, y = -1.8, parse = TRUE, label = "T['surface'] <  T['deep']") +
  Mytheme(size_labs = 12) +
  theme(panel.grid.major.x = element_blank(), panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())
print(temp_delta_plt)
```

@@

```{r, echo=FALSE, message= FALSE, warning=FALSE}
# we look at density at all depths
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, temp_insitu_ctd, depth_group, pressure_insitu_ctd) %>%
  dplyr::filter(depth_group != "NA") %>% 
  dplyr::mutate(depth_group = factor(depth_group, levels = depth_labels)) %>% 
  dplyr::mutate(density = seacarb::rho(S = sal_insitu_ctd, T = temp_insitu_ctd, P = pressure_insitu_ctd)) %>%
p <- tidyr::drop_na(tmp) %>% 
  ggplot(aes(x=as.factor(depth_group), y = sal_insitu_ctd)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "Density vs depth", x = "Depth (m)", y= "Density") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$density ~ tmp$depth_group * tmp$montht))

# we look at density 0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, sal_insitu_ctd, temp_insitu_ctd, depth_group, pressure_insitu_ctd) %>%
  dplyr::filter(depth_group != "NA") %>% 
  dplyr::mutate(depth_group2 = as.factor(case_when(depth_group == '0-2' | depth_group == '2-4' ~ '0-4', 
                                      depth_group == '>8' ~ '8-12')), 
                density = seacarb::rho(S = sal_insitu_ctd, T = temp_insitu_ctd, 
                                       P = pressure_insitu_ctd)) %>% 
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>% 
  dplyr::group_by(depth_group2, montht)

@@

```{r, echo=FALSE, message= FALSE, warning=FALSE}
# we look at pHint_tot_sf at all depths
tmp <- d_all %>% 
  dplyr::select(montht, pHint_tot_sf, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  dplyr::mutate(depth_group = factor(depth_group, levels = depth_labels))
p <- tmp %>% 
  ggplot(aes(x = as.factor(depth_group), y = pHint_tot_sf)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pHint_tot_sf vs depth", x = "Depth (m)", y= "pH (units)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pHint_tot_sf ~ tmp$depth_group * tmp$montht))

@@

tmp_delta <- tmp %>%
  dplyr::summarise(pH_tot_sf_med = median(pHint_tot_sf, na.rm = TRUE)) %>%
  tidyr::pivot_wider(names_from = depth_group2, values_from = pH_tot_sf_med) %>%
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>%   dplyr::mutate(pH_tot_sf_med_delta = median0_4m - median8_12m)
kable(tmp_delta, digits = 3, caption = "Median seafet pH with (median0_4m - median8_12m)")

@@

# we look at pHint_tot_sf normalized at 4°C at all depths
tmp <- d_all %>% 
  dplyr::select(montht, pHint_tot_sf_temp_4, depth_group) %>%
  dplyr::filter(depth_group != "NA") %>% 
  dplyr::mutate(depth_group = factor(depth_group, levels = depth_labels))
p <- tmp %>% 
  drop_na() %>% 
  ggplot(aes(x=as.factor(depth_group), y = pHint_tot_sf_temp_4)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pHint_tot_sf at 4°C vs depth", x = "Depth (m)", y= "pH at 4°C (units)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pHint_tot_sf_temp_4 ~ tmp$depth_group * tmp$montht))

# we look at pHint_tot_sf normalized at 4°C  0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, pHint_tot_sf_temp_4, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(case_when(depth_group == '0-2' | depth_group == '2-4' ~ '0-4',
                                                   depth_group == '>8' ~ '8-12'))) %>% 
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>% 
summary(aov(tmp$pHint_tot_sf_temp_4 ~ tmp$depth_group2 * tmp$montht))

tmp_delta <- tmp %>% 
  dplyr::summarise(pH_tot_sf_med = median(pHint_tot_sf_temp_4, na.rm = TRUE)) %>% 
  tidyr::pivot_wider(names_from = depth_group2, values_from = pH_tot_sf_med) %>% 
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>%   dplyr::mutate(pH_tot_sf_med_delta = median0_4m - median8_12m)
kable(tmp_delta, digits = 3, caption = "Median seafet pH at 4°C with (median0_4m - median8_12m)")

@@

# we look at pHint_tot_sf normalized at 4°C at all depths
tmp <- d_all %>% 
  dplyr::select(montht, pHint_tot_sf_temp_4, depth_group) %>%
  dplyr::filter(depth_group != "NA") %>% 
  dplyr::mutate(depth_group = factor(depth_group, levels = depth_labels))
p <- tmp %>% 
  tidyr::drop_na() %>% 
  ggplot(aes(x=as.factor(depth_group), y = pHint_tot_sf_temp_4)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pHint_tot_sf at 4°C vs depth", x = "Depth (m)", y= "pH at 4°C (units)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pHint_tot_sf_temp_4 ~ tmp$depth_group * tmp$montht))

# we look at pHint_tot_sf normalized at 4°C  0 to 4 m vs more than 8 m
tmp <- d_all %>% 
  dplyr::select(montht, pHint_tot_sf_temp_4, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(case_when(depth_group == '0-2' | depth_group == '2-4' ~ '0-4',
                                                   depth_group == '>8' ~ '8-12'))) %>% 
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12') %>% 
summary(aov(tmp$pHint_tot_sf_temp_4 ~ tmp$depth_group2 * tmp$montht))

@@

```{r, echo=FALSE, message= FALSE, warning=FALSE}
# we look at pCO2 NOT normalized at 4°C at all depths
# calculate pCO2 from pH and salinity-derived AT
tmp <- d_all %>%
  dplyr::select(datetime,
                montht,
                depth_group,
                depth,
                pHint_tot_sf,
                sal_insitu_ctd,
                temp_insitu_ctd) %>%
  drop_na() %>%
  dplyr::mutate(
    at_remos = (intercept_s_at + slope_s_at * sal_insitu_ctd) * 1e-6,
    pco2_remos_ctd_temp = seacarb::carb(
      flag = 8,
      var1 = pHint_tot_sf,
      var2 = at_remos,
      S = sal_insitu_ctd,
      T = temp_insitu_ctd,
      Patm = 1,
      P = depth / 10,
      Pt = 0,
      Sit = 0,
      k1k2 = "l",
      kf = "dg",
      ks = "d",
      pHscale = "T",
      b = "u74"
    )$pCO2
  )
# add at_remos and pco2_remos_ctd_temp to d_all
d_all <-
  dplyr::left_join(d_all, dplyr::select(tmp, datetime, at_remos, pco2_remos_ctd_temp), by = "datetime") %>%
  dplyr::select(sort(current_vars()))
tmp <- tmp %>% 
  dplyr::select(montht, pco2_remos_ctd_temp, depth_group) %>%
  dplyr::filter(depth_group != "NA") %>% 
  dplyr::mutate(depth_group = factor(depth_group, levels = depth_labels))
p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = pco2_remos_ctd_temp)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pco2_sf vs depth", x = "Depth (m)", y= "fCO2 (uatm)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pco2_remos_ctd_temp ~ tmp$depth_group * tmp$montht))

# we look at pCO2 NOT normalized at 4°C 0 to 4 m vs more than 8 m
tmp <- tmp %>%
  dplyr::select(montht, pco2_remos_ctd_temp, depth_group) %>%
  dplyr::mutate(depth_group2 = as.factor(
    case_when(
      depth_group == '0-2' | depth_group == '2-4' ~ '0-4',
      depth_group == '>8' ~ '8-12'
    )
  )) %>%
  dplyr::filter(depth_group2 == '0-4' | depth_group2 == '8-12')
summary(aov(tmp$pco2_remos_ctd_temp ~ tmp$depth_group2 * tmp$montht))

tmp_delta <- tmp %>%
  dplyr::summarise(pco2_remos_ctd_temp_med = median(pco2_remos_ctd_temp, na.rm = TRUE)) %>%
  tidyr::pivot_wider(names_from = depth_group2, values_from = pco2_remos_ctd_temp_med) %>%
  dplyr::rename(median0_4m = "0-4",
                median8_12m = "8-12") %>%
  dplyr::mutate(pco2_remos_ctd_temp_med_delta = median0_4m - median8_12m) %>%
  # this is to bring June back...
  rbind(
    data.frame(
      montht = "Jun",
      median8_12m = NA,
      median0_4m = NA,
      pco2_remos_ctd_temp_med_delta = NA
    )
  ) %>%
  dplyr::mutate(montht = factor(montht, levels = month.abb)) %>%
  dplyr::arrange(montht)
kable(tmp_delta, digits = 3, caption = "Median REMOS pCO2 (median0_4m - median8_12m)")

@@

tmp <- tmp %>% 
  dplyr::select(montht, pco2_remos_4, depth_group) %>%
  tidyr::drop_na() %>% 
  dplyr::filter(depth_group != "NA") %>% 
  dplyr::mutate(depth_group = factor(depth_group, levels = depth_labels))

p <- tmp %>% 
  ggplot(aes(x=as.factor(depth_group), y = pco2_remos_4)) +
  geom_boxplot(notch=FALSE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(title = "pco2_remos at 4°C vs depth", x = "Depth (m)", y= "pCO2 (uatm)") +
  scale_x_discrete(labels = depth_labels) +
  facet_wrap(vars(montht), scale = "free_y")
print(p)
summary(aov(tmp$pco2_remos_4 ~ tmp$depth_group * tmp$montht))
